\documentclass[12pt]{article}

% Packages
\usepackage[margin=6em]{geometry} % 1 cm = 2.84528 em
\usepackage[colorlinks=true,linkcolor=red,urlcolor=blue]{hyperref}
\usepackage{lipsum}

% Paragraphs
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

% Includes
\input{antonxue-lib}

\let\H\undefined
\newcommand{\H}{\mathcal{H}}

\let\F\undefined
\newcommand{\F}{\mathcal{F}}

% Author
\author{}
\title{Discrete Fourier Transforms for Non-linear Real Arithmetics}
% \date{\today}
\date{}

% Document
\begin{document}
\maketitle

\subsection{The Problem}
We first consider the satisfiability of first-order logic formulae over
non-linear real arithmetics:

\begin{align*}
  \varphi
    = \bigwedge_{i = 1}^{N} \parens{\bigvee_{j = 1}^{M}
        f_{i, j}\parens{\vec{x}} \leq \varepsilon_{i, j}}
  \qquad
  \vec{x} \in K \subseteq \R^n, \enskip
  \type{f_{i, j}}{K}{\R}, \enskip
  \varepsilon_{i, j} \geq 0
\end{align*}

The quesiton of satisfiability, however, is undecidable:
we cannot yield algorithms that, given $\varphi$, is able to find some
$\vec{x}$ if there exists a satisfying configuration.
We thus seek to approximate solutions, which throws away soundness for
feasibility.

However, we are interested in still another class of problems:
program synthesis.
Given a pre-condition $P$ that we assume at the start of program execution,
and post-condition $Q$, does there exists some program $S$ that is able to
transform $P$ into $Q$?
When $P$ and $Q$ are formulae over linear integer arithmetics, for instance,
the corresponding program is akin to a sequence of variable assignments
that correspond to a linear transformation.
We never said these programs have to be fancy, did we?


\subsection{Background}
\subsubsection*{The Fourier Transform}
Consider the separable Hilbert space $\H = L^2 \parens{\brackets{0, 1}}$
endowed with the Fourier basis $\braces{e_n}_{n \in \Z}$, where we define
$e_n \parens{x} = e^{i 2\pi n x}$.
A Fourier transform $\type{\F_n}{\H}{\C}$ is a linear functional that can
be expressed as an inner product:

\begin{align*}
  \F_n\brackets{f}
    = \inner{f, e_n}
    = \int_0^1 f\parens{x} \overline{e_n \parens{x}} dx
    = \int_0^1 f\parens{x} e^{-i 2 \pi n x} dx
\end{align*}

We further claim that the Fourier transform is continuous with
norm $1$:

\begin{align*}
  \norm{F_n}_{\adj{\H}}
    = \sup_{\norm{f}_{\H} = 1}
        \abs{\int_0^1 f\parens{x} \overline{e_n \parens{x}} dx}
    \leq \parens{\int_0^1 \abs{f\parens{x}}^2 dx}^{1/2}
              \parens{\int_0^1 \abs{e_n \parens{x}}^2 dx}^{1/2}
    = \norm{f}_{\H} \cdot \norm{e_n}_{H}
\end{align*}

The inequality is sharp when $f = e_{n}$.
We recall that the partial Fourier series $\type{S_n}{\H}{\H}$ is a
(finite dimensional) compact operator:

\begin{align*}
  S_N \brackets{f} \parens{x}
    = \sum_{n = -N}^{N} \inner{f, e_n} e_n \parens{x}
    = \sum_{n = -N}^{N} e_n \parens{x} \int_0^1 f(y) e^{- i 2 \pi n y} dy
\end{align*}

Note that this is an orthogonal projection.
Consider some $m$ such that
$\abs{m} > N$:

\begin{align*}
  \inner{e_m, S_n}
    = \inner{e_m, \sum_{n = -N}^{N} \inner{f, e_n} e_n}
    = \sum_{n = -N}^{N} \inner{f, e_n} \cdot \inner{e_m, e_n}
    = \sum_{n = -N}^{N} \inner{f, e_n} \cdot 0
    = 0
\end{align*}

Furthermore, we omit the proof, but do note that this converges in norm
for any $f \in \mathcal{H}$:
\begin{align*}
  \lim_{N \to \infty} \norm{f - S_N \brackets{f}}_{\H}
    = \lim_{N \to \infty}
      \abs{f\parens{x} -
            \sum_{n = -N}^{N} \int_0^1 f\parens{x} e^{-i 2\pi nx} dx}
    \to 0
\end{align*}

We also remark that the class of continuous functions over
compact support are dense in $L^2$.
This motivates us to consider the problem of non-linear real arithmetics
over compact (perhaps convex?) domains.


\subsubsection{Multi-Dimensional Fourier Transform}
We now consider the Fourier transform when we consider $n$ dimensions.
For $\H = L^2 \parens{\brackets{0, 1}^n}$.
We define as the orhonormal basis:

\begin{align*}
  \phi_{d_1, \ldots, d_n} \parens{x_1, \ldots, x_n}
    = \prod_{i = 1}^{n} e_{d_i} \parens{x_i}
    = e^{i 2 \pi \parens{d_1 x_1 + \cdots + d_n x_n}}
    \qquad
    d_1, \ldots, d_n \in \Z
\end{align*}

We likewise define the
Fourier transform $\type{F_{i_1, \ldots, i_n}}{\R^n}{\C}$ as a
linear functional in the form of an inner product,
with $\vec{x} \in \R^n$:

\begin{align*}
  \mathcal{F}_{i_1, \ldots, i_n} \brackets{f} \parens{\vec{x}}
    = \inner{f, \phi_{i_1, \ldots, i_n}}
    = \int_{\brackets{0, 1}^n}
      f\parens{\vec{x}} \overline{\phi_{i_1, \ldots, i_n} \parens{\vec{x}}} d \vec{x}
\end{align*}

A Fourier series for $N_1, \ldots, N_n$:

\begin{align*}
  S_{N_1, \ldots, N_n} \brackets{f} \parens{\vec{x}}
    = \sum_{i_1 = -N_1}^{N_1} \cdots \sum_{i_n = -N_n}^{N_n}
        \inner{f, \phi_{i_1, \ldots, i_n}} \phi_{i_1, \ldots, i_n} \vec{x}
\end{align*}

I am sure a proof of convergence in $L^2$ norm for this exists somewhere.

\subsubsection{Discrete Fourier Transform}



\subsection{The Problem}


\textbf{We may approximate the precondition as a matrix of linear inequalities
over the dual space of the domain, and likewise for the codomain.
This therefore becomes an LP problem with respect to the dual spaces, and
all that's left is figuring out how fast the L2 norms converge}


\end{document}

